A comparison of classification algorithms applied to the Adult Database

### 2009-12-10 report/final.tex
updates.

In Weka, the Random Forest algorithm can be parametrized by the number
of trees to build \textbf{$<$-I integer$>$}. Tables \ref{tbl:results_random_forest_discr}
and \ref{tbl:results_random_forest_raw} show results for
a varying number of trees, for discretized data and non-discretized data, respectively.

The classifier built with non-discretized data can easily reach a satisfactory
error rate with less trees ($\approx$4 against $\approx$40).

For both cases, the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
when comparing to the test error rate.

@@

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    -I 2 & 10.52 & 34.03 & 0m25.861s \\ \hline
    -I 3 & 9.83 & 34.00 & 0m26.318s \\ \hline
    -I 4 & 9.69 & 30.91 & 0m26.154s \\ \hline
    -I 5 & 9.50 & 29.62 & 0m27.108s \\ \hline
    -I 7 & 9.32 & 28.03 & 0m27.158s \\ \hline
    -I 10 & 9.24 & 29.80 & 0m28.107s \\ \hline
    -I 15 & 9.18 & 28.80 & 0m30.716s \\ \hline
    -I 20 & 9.15 & 28.09 & 0m36.340s \\ \hline
    -I 40 & 9.15 & 23.43 & 0m39.687s \\ \hline
    -I 60 & 9.15 & 23.65 & 0m49.443s \\ \hline
    -I 80 & 9.15 & 22.92 & 0m57.652s \\ \hline
    -I 100 & 9.15 & 22.83 & 1m28.024s \\ \hline
    -I 400 & 9.15 & 22.71 & 4m58.842s \\ \hline
  \end{tabular}
  \caption{Results for Random Forest using discretized data.}
  \label{tbl:results_random_forest_discr}
  \end{center}
\end{table}

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    -I 2 & 5.72 & 25.70 & 0m4.882s \\ \hline
    -I 3 & 3.89 & 24.58 & 0m5.588s \\ \hline
    -I 4 & 3.81 & 23.40 & 0m6.086s \\ \hline
    -I 5 & 3.20 & 23.43 & 0m6.839s \\ \hline
    -I 7 & 2.86 & 21.01 & 0m8.190s \\ \hline
    -I 10 & 2.69 & 21.15 & 0m10.162s \\ \hline
    -I 15 & 2.52 & 21.01 & 0m13.597s \\ \hline
    -I 20 & 2.46 & 20.83 & 0m16.970s \\ \hline
    -I 40 & 2.44 & 21.26 & 0m30.488s \\ \hline
    -I 60 & 2.44 & 21.67 & 0m46.093s \\ \hline
    -I 80 & 2.44 & 21.70 & 1m1.347s \\ \hline
    -I 100 & 2.44 & 21.75 & 1m15.592s \\ \hline
    -I 400 & 2.44 & 21.62 & 7m51.624s \\ \hline
  \end{tabular}
  \caption{Results for Random Forest using non-discretized data.}
  \label{tbl:results_random_forest_raw}
  \end{center}
\end{table}

@@

    -M 5 -N 5 & 13.94 & 25.41 & 2m43.363s & 123 & 62 \\ \hline
    -M 5 -N 10 & 13.82 & 25.34 & 4m54.547s & 145 & 73 \\ \hline
    -M 5 -N 15 & 13.94 & 25.41 & 7m9.537s & 123 & 62 \\ \hline
    -M 5 -N 20 & 13.88 & 25.39 & 9m18.095s & 133 & 67 \\ \hline
  \end{tabular}
  \caption{Results for CART with discretized data.}
  \label{tbl:results_cart_discr}
  \end{center}
\end{table}

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
    -M 2 -N 5 & 12.68 & 22.26 & 2m34.703s & 157 & 79 \\ \hline
    -M 2 -N 10 & 12.72 & 22.26 & 4m49.261s & 151 & 76 \\ \hline
    -M 2 -N 15 & 12.22 & 21.34 & 7m4.560s & 237 & 119 \\ \hline
    -M 2 -N 20 & 11.75 & 21.70 & 9m14.989s & 341 & 171 \\ \hline

    -M 3 -N 5 & 12.68 & 22.26 & 2m21.439s & 157 & 79 \\ \hline
    -M 3 -N 10 & 12.72 & 22.26 & 4m35.048s & 151 & 76 \\ \hline
    -M 3 -N 15 & 12.22 & 21.34 & 6m45.982s & 237 & 119 \\ \hline
    -M 3 -N 20 & 11.84 & 21.70 & 7m44.483s & 321 & 161 \\ \hline

    -M 5 -N 5 & 12.72 & 22.26 & 2m18.255s & 151 & 76 \\ \hline
    -M 5 -N 10 & 12.60 & 22.78 & 4m28.169s & 171 & 86 \\ \hline
    -M 5 -N 15 & 13.29 & 22.01 & 6m45.657s & 83 & 42 \\ \hline
    -M 5 -N 20 & 12.22 & 21.43 & 8m53.232s & 259 & 130 \\ \hline
  \end{tabular}
  \caption{Results for CART with non-discretized data.}
  \label{tbl:results_cart_raw}
  \end{center}
\end{table}

CART results are displayed in tables \ref{tbl:results_cart_discr} and
\ref{tbl:results_cart_raw}.
From these results we can see that there isn't a lot of error rate variation
for each type of data.
The algorithm also performs worse compared to C4.5 or ADTree
and takes a much longer time to build the classifier.

>>>>

In Weka, the Random Forest algorithm can be parametrized by the number
of trees to build \textbf{$<$-I integer$>$}. Table \ref{tbl:results_random_forest} shows results for
a varying number of trees.

From the table we conclude that we need a certain amount of trees ($\approx$80) to
reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
the error rate for the train set is low ($\approx$9\%), which indicates over-fitting
when comparing the difference to the test error rate.

@@

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    -I 2 & 10.52 & 34.03 & 0m25.861s \\ \hline
    -I 3 & 9.83 & 34.00 & 0m26.318s \\ \hline
    -I 4 & 9.69 & 30.91 & 0m26.154s \\ \hline
    -I 5 & 9.50 & 29.62 & 0m27.108s \\ \hline
    -I 7 & 9.32 & 28.03 & 0m27.158s \\ \hline
    -I 10 & 9.24 & 29.80 & 0m28.107s \\ \hline
    -I 15 & 9.18 & 28.80 & 0m30.716s \\ \hline
    -I 20 & 9.15 & 28.09 & 0m36.340s \\ \hline
    -I 40 & 9.15 & 23.43 & 0m39.687s \\ \hline
    -I 60 & 9.15 & 23.65 & 0m49.443s \\ \hline
    -I 80 & 9.15 & 22.92 & 0m57.652s \\ \hline
    -I 100 & 9.15 & 22.83 & 1m28.024s \\ \hline
    -I 400 & 9.15 & 22.71 & 4m58.842s \\ \hline
  \end{tabular}
  \caption{Results for Random Forest.}
  \label{tbl:results_random_forest}
  \end{center}
\end{table}

@@

    -M 5 -N 5 & 13.94 & 25.41 & 2m43.363s & 123 & 62 \\ \hline
    -M 5 -N 10 & 13.82 & 25.34 & 4m54.547s & 145 & 73 \\ \hline
    -M 5 -N 15 & 13.94 & 25.41 & 7m9.537s & 123 & 62 \\ \hline
    -M 5 -N 20 & 13.88 & 25.39 & 9m18.095s & 133 & 67 \\ \hline
  \end{tabular}
  \caption{Results for CART.}
  \label{tbl:results_cart}
  \end{center}
\end{table}

CART results are displayed in Table \ref{tbl:results_cart}.
From these results we can see that there isn't a lot of variation in the results.
The algorithm also performs badly when compared to C4.5 or ADTree.
